|

### 2024-03-26 manuscript/manuscript.qmd
update

  affiliations:
    name: Florida International University

@@

  affiliations:
    name: Texas Tech University

@@

  affiliations:
    name: Oregon Health & Science University

@@

  affiliations:
    name: University of Missouri

@@

  affiliations:
    name: University of California San Diego

@@

  affiliations:
    name: University of Pittsburgh

@@

  affiliations:
    name: Virginia Commonwealth University

@@

  affiliations:
    name: University of Michigan

@@

  affiliations:
    name: University of Oxford

@@

  affiliations:
    name: University of Pennsylvania

@@

  affiliations:
    name: Laureate Institute for Brain Research

@@

abstract: |
  The Adolescent Brain Cognitive Development&reg; (ABCD) Study provides a unique opportunity for researchers to investigate developmental processes in a large, diverse cohort of youths, aged 9-10 at baseline and assessed annually for 10 years. Given the size and complexity of the ABCD Study, researchers analyzing its data will encounter a myriad of methodological and analytical considerations. This review provides an examination of key concepts and techniques related to longitudinal analyses of the ABCD Study data.  We discuss the importance of longitudinal data, focusing on the types of inferences that are possible when one assesses individuals across multiple time points, including: 1) characterization of the factors associated with variation in developmental trajectories; 2) assessment of how level and timing of exposures may impact subsequent development; 3) quantification of how variation in developmental domains may be associated with outcomes, including mediation models and reciprocal relationships. We emphasize the importance of selecting appropriate statistical models to address these research questions, e.g., accounting for correlation in repeated measurements, employing linear or non-linear models as indicated by the data, and using link functions that adequately model outcome distributions. By presenting the advantages and potential challenges of longitudinal analyses in the ABCD Study, this review seeks to equip researchers with foundational knowledge and tools to make informed decisions as they navigate and effectively analyze and interpret the multi-dimensional longitudinal data currently available.

@@

format:
  ieee-typst: default
  html:
    page-layout: article
    toc: true
    toc-depth: 1
    toc-expand: 2
    toc-title: Table of Contents
    code-tools:
      source: true
      toggle: false
      caption: none
    grid:
      sidebar-width: 350px
      margin-width: 20px
bibliography: references.bib

>>>>

  #affiliations:
    #name: Florida International University

@@

  #affiliations:
    #name: Texas Tech University

@@

  #affiliations:
    #name: Oregon Health & Science University

@@

  #affiliations:
    #name: University of Missouri

@@

  #affiliations:
    #name: University of California San Diego

@@

  #affiliations:
    #name: University of Pittsburgh

@@

  #affiliations:
    #name: Virginia Commonwealth University

@@

  #affiliations:
    #name: University of Michigan

@@

  #affiliations:
    #name: University of Oxford

@@

  #affiliations:
    #name: University of Pennsylvania

@@

  #affiliations:
    #name: Laureate Institute for Brain Research

@@

abstract: |
  The Adolescent Brain Cognitive Development&reg; (ABCD) Study provides a unique opportunity to investigate developmental processes in a large, diverse cohort of youths, aged approximately 9-10 at baseline and assessed annually for 10 years. Given the size and complexity of the ABCD Study, researchers analyzing its data will encounter a myriad of methodological and analytical considerations. This review provides an examination of key concepts and techniques related to longitudinal analyses of the ABCD Study data, including: 1) characterization of the factors associated with variation in developmental trajectories; 2) assessment of how level and timing of exposures may impact subsequent development; 3) quantification of how variation in developmental domains may be associated with outcomes, including mediation models and reciprocal relationships. We emphasize the importance of selecting appropriate statistical models to address these research questions. By presenting the advantages and potential challenges of longitudinal analyses in the ABCD Study, this review seeks to equip researchers with foundational knowledge and tools to make informed decisions as they navigate and effectively analyze and interpret the multi-dimensional longitudinal data currently available.

@@

format:
  docx: default
  ieee-typst: default
  html:
    page-layout: article
    toc: true
    toc-depth: 1
    toc-expand: 2
    toc-title: Table of Contents
    code-tools:
      source: true
      toggle: false
      caption: none
    grid:
      sidebar-width: 350px
      margin-width: 20px
bibliography: references.bib
